---
title: "Plants-and-Bees"
author: "JT Miller"
date: '2022-07-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Necessary packages
```{r}
library(tidyverse)

```
### Load in the Globi-Interaction database 
```{r}
globi_bee_data <- read.csv("/home/jt-miller/Globi-Bees-Plant-Interactions/global-bee-interaction-data/all_bee_data.csv", quote = "") # Disable quoting might be necessary for this dataset due to EOF warning. Solution found here: https://stackoverflow.com/questions/17414776/read-csv-warning-eof-within-quoted-string-prevents-complete-reading-of-file
```

### Data Exploration
Lets first figure out what we're dealing with
```{r}
names(globi_bee_data)
```
Appears that either "sourceBodyPartName" or "targetBodyPartName" will be important to identify whether it was actually a flower visitation interaction as well as "interactionTypeName". There is also "decimalLatitude" and "decimalLongitude" so we should be able to map it down to only include interactions that occur in California. 
