---
title: "Climate-Comparisons"
author: "JT Miller"
date: "2022-08-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### This md is an analysis on climate in the Extended California Ecoregions

```{r}
library(tidyverse)
```


First, bring in the data. Data was pulled from climateNA using the years 1961-1990 (These years are before the large climatic shift, therefore we're going to start with this as a baseline for comparison)
```{r}
ecoregion_grid <- read.csv("/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-grid-pull2_Normal_1961_1990Y.csv")
```

Lets take a look at mean annual temperature using boxplots for the ecoregions
```{r}

# Since Simpsons Paradox is a thing, lets try to visualize our data a bit 







ggplot(data = ecoregion_grid, aes(x = ecoRegion, y = MAT, fill = ecoRegion)) +
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_blank())
  
```
And lets compare that to where the bees are found using the climateNA pulls on their occurrences
```{r}
bee_grid <- read.csv("/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-extended-regions-bee-pull2_Normal_1961_1990Y.csv")

ggplot(data = bee_grid, aes(x = ecoRegion, y = MAT, fill = ecoRegion)) +
  geom_boxplot() +
  theme_bw() + 
  theme(axis.text.x = element_blank())

bee_grid %>% 
  group_by(ecoRegion) %>% 
  summarise(mean(MAT), mean(MWMT), mean(MCMT), mean(TD), mean(DD5), mean(MAP))


# Comparison 

ecoregion_grid %>% 
  group_by(ecoRegion) %>% 
  summarise(mean(MAT), mean(MWMT), mean(MCMT), mean(TD), mean(DD5), mean(MAP))

```
Histogram Comparisons 
```{r}

shap_oak_woodlands_grid <- subset(ecoregion_grid, ecoregion_grid$ecoRegion == "Cali_Coastal_Sage_Chap_Oak_Woodlands")

Chap_bees <- bee_grid %>% 
  filter(ecoRegion =="Cali_Coastal_Sage_Chap_Oak_Woodlands")

ggplot(data = shap_oak_woodlands_grid, aes(x = MAT)) +
  geom_histogram()


ggplot(data = Chap_bees, aes(x = MAT)) +
  geom_histogram()
```

### Creating a Climograph using the gridded regions with season data. 

bring in the data 
```{r}
regions_climate <- read.csv("/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-extended-regions-grid-pull2_Normal_ALL_1961_1990MSY.csv")

library(tidyverse)
# Find possible NAs 
regions_climate_NAs <- regions_climate %>% 
  filter(Tmax01 == -9999) # There are 23 occurrence pts present with listed NAs, lets drop them. 

regions_climate_drop <- regions_climate %>% 
  filter(!(Tmax01 == -9999))
```

Take Averages of variables of interest. 
```{r}
regions_climate_avgs <- regions_climate_drop %>% 
  group_by(ecoRegion) %>%  # Group the following actions by ecoRegion
  mutate(across(contains("Tave"), mean, # Go across any column that matches the string Tave and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_Tave"), round, 1)) %>%  # Round the columns to one decimal point, note this isn't sig figs but the way climateNA had their data set up. Probably check into this later. 
  mutate(across(contains("PPT"), mean, # Do the same for precipitation
          .names = "{fn}_{col}")) %>% 
  mutate(across(contains("1_PPT"), round, 0)) %>% 
  mutate(across(contains("RAD"), mean, # Go across any column that matches the string RAD and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_RAD"), round, 1)) %>%
  mutate(across(contains("DD5"), mean, # Go across any column that matches the string DD5 and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_DD5"), round, 0)) %>%
  mutate(across(contains("RH"), mean, # Go across any column that matches the string RH and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_RH"), round, 0)) %>%
  mutate(across(contains("CMI"), mean, # Go across any column that matches the string CMI and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_CMI"), round, 0)) %>% 
  mutate(across(contains("Eref"), mean, # Go across any column that matches the string RH and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_Eref"), round, 0)) %>% 
  mutate(across(contains("Tmax"), mean, # Go across any column that matches the string Tave and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_Tmax"), round, 1)) %>% 
  mutate(across(contains("Tmin"), mean, # Go across any column that matches the string Tave and give a mean
         .names = "{fn}_{col}")) %>% # Rename the columns
  mutate(across(contains("1_Tmin"), round, 1))
  
  

regions_avgs <- rename_with(regions_climate_avgs, ~ gsub("1_", "avg_", .x, fixed = TRUE))

regions_avgs_d <- distinct(regions_avgs, ecoRegion, .keep_all = TRUE)

regions_avgs_less <- regions_avgs_d %>% 
  select(ecoRegion, contains("avg_"))

### Reorganize this df for later use ###

regions_avg_temp <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_Tave"), 
    names_to = "temp_month",
    values_to = "avg_temp_value") %>% 
  select(ecoRegion, temp_month, avg_temp_value)

regions_avg_PPT <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_PPT"), 
    names_to = "PPT_month",
    values_to = "avg_PPT_value") %>% 
  select(ecoRegion, PPT_month, avg_PPT_value)

regions_avg_Rad <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_Rad"), 
    names_to = "Rad_month",
    values_to = "avg_Rad_value") %>% 
  select(ecoRegion, Rad_month, avg_Rad_value)

regions_avg_DD5 <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_DD5"), 
    names_to = "DD5_month",
    values_to = "avg_DD5_value") %>% 
  select(ecoRegion, DD5_month, avg_DD5_value)

regions_avg_RH <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_RH"), 
    names_to = "RH_month",
    values_to = "avg_RH_value") %>% 
  select(ecoRegion, RH_month, avg_RH_value)

regions_avg_CMI <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_CMI"), 
    names_to = "CMI_month",
    values_to = "avg_CMI_value") %>% 
  select(ecoRegion, CMI_month, avg_CMI_value)

regions_avg_Eref <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_Eref"), 
    names_to = "Eref_month",
    values_to = "avg_Eref_value") %>% 
  select(ecoRegion, Eref_month, avg_Eref_value)

regions_avg_Tmax <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_Tmax"), 
    names_to = "Tmax_month",
    values_to = "avg_Tmax_value") %>% 
  select(ecoRegion, Tmax_month, avg_Tmax_value)

regions_avg_Tmin <- regions_avgs_less %>% 
  pivot_longer(
    cols = contains("avg_Tmin"), 
    names_to = "Tmin_month",
    values_to = "avg_Tmin_value") %>% 
  select(ecoRegion, Tmin_month, avg_Tmin_value)

regions_temp_percipitation <- cbind(regions_avg_PPT, regions_avg_temp, regions_avg_Tmax, regions_avg_Tmin)

month_v <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec", "Winter", "Spring", "Summer", "Fall")

regions_df <- regions_temp_percipitation %>% 
  select(ecoRegion...1, avg_PPT_value, avg_temp_value,avg_Tmax_value,avg_Tmin_value) %>% 
  add_column(months = rep(month_v, 12)) %>% 
  rename(ecoRegion = ecoRegion...1)
###
 



list_1 <- lapply(1:length(unique(regions_avgs_d$ecoRegion)),
                function(i)filter(regions_avgs_d, ecoRegion == unique(regions_avgs_d$ecoRegion)[i]))

df_names <- c(regions_avgs_d$ecoRegion)

split_df <- function(list, df_names){
  for(i in 1:length(list)){
   
    assign(paste0(df_names[i]), list[[i]], envir = .GlobalEnv)
  
}
  }


list_1


```

```{r}
### Split up the dataframes using our new fxn. 
split_df(list_1, df_names)

```

Now to make a climogram/climograph. The goal is to create a full climate analysis for each ecoregion. 
requirements:
* X-axis should be the 12 months (& variation with the Seasons) 
* Y-axis #1 should be percipitation 
* Y-axis #2 should be Temperature
* Somehow include Max Temp (Red dot?)
* Somehow include Coldest Temp (Blue dot?)
* Add Solar Radiation MJ m^-2 d^-1 (?)
```{r}
# Example code for a climatic graph using ggplot2
df <- as.data.frame(c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
colnames(df) <- c("month")
df$month <- factor(df$month, levels = month.abb)
df$celsius <- c(-26.0, -24.5, -18.9, -9.8, -1.0, 7.0, 12.7, 12.3, 6.4, -1.2, -12.7, -21.9)
df$prec_mm <- c(18.7, 16.6, 18.1, 23.6, 30.0, 44.2, 59.8, 69.4, 69.9, 48.4, 35.5, 18.4)


df2 <- as.data.frame(c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
colnames(df2) <- c("month")
df2$month <- factor(df$month, levels = month.abb)
df2$celsius <- c(Coast_Range[,271:282])
df2$prec_mm <- c(Coast_Range[,287:298])
df2$max_celsius <- c(Coast_Range[,387:398])
df2$min_celsius <- c(Coast_Range[,403:414])

df2$celsius <- unlist(df2$celsius)
df2$prec_mm<- unlist(df2$prec_mm)
df2$max_celsius <- unlist(df2$max_celsius)
df2$min_celsius <- unlist(df2$min_celsius)

ggplot(data = df, mapping = aes(x = month, y = prec_mm, group = 1)) + 
  geom_bar(stat = "identity", color="blue", fill="blue", width = 0.5) + 
  geom_line(mapping = aes(y = celsius+30), color="red", size=1.5) + # Scale data to match desired scale
  scale_y_continuous("Precipitation [mm]", 
                     sec.axis = sec_axis(~ . -30, name = "Temperature [째C]") # Reverse transformation to match data
  )

thinned <- floor(seq(from=1,to=dim(df)[1],length=70))

ggplot(data = df, mapping = aes(x = month, y = prec_mm, group = 1)) + 
  geom_bar(stat = "identity", color="blue", fill="blue", width = 0.5) + 
  geom_line(mapping = aes(y = celsius+30), color="red", size=1.5) + # Scale data to match desired scale
  geom_point(data=df[thinned,],aes(x=month,y=celsius+30)) +
  scale_y_continuous("Precipitation [mm]", 
                     sec.axis = sec_axis(~ . -30, name = "Temperature [째C]") # Reverse transformation to match data
  )

ggplot(data = df2, mapping = aes(x = month, y = prec_mm, group = 1)) + 
  geom_bar(stat = "identity", color="blue", fill="blue", width = 0.5) + 
  geom_line(mapping = aes(y = celsius), color="red", size=1.5) + # Scale data to match desired scale
  scale_y_continuous("Precipitation [mm]", 
                     sec.axis = sec_axis(~ . *10, name = "Temperature [째C]") # Reverse transformation to match data
  )
```
### Trying a slightly different method 
```{r}
### This is what I like. 
df2 %>% 
  ggplot(aes(x=month, group=1)) + 
  geom_bar(aes(y=prec_mm/10, col = "Percipitation  [mm]"), stat = "identity", fill = "lightblue") +
  geom_line(aes(y=celsius, col="Avg. Temp [C]"), size = 1) + 
  geom_line(aes(y=max_celsius, col = "Max. Avg. Temp [C]")) +
  geom_line(aes(y=min_celsius, col = "Min. Avg.  Temp [C]")) +
  geom_point(aes(y=celsius)) +
  theme_bw() +
  scale_color_manual(values=c("darkorange", "red" ,"blue", "black")) +
  scale_y_continuous(sec.axis = sec_axis(~.*10, name = "Percipitation [mm]")) +
  labs(x="month", y="Temperature [C]")
```
```{r}
# Now lets come up with a fancy way of doing all of that in a facet grid.
regions_df_monthly <- regions_df %>% 
  filter(!(months %in% c("Winter", "Spring", "Summer", "Fall")))

regions_df_monthly$months <- as.character(regions_df_monthly$months)

regions_df_monthly$months <- factor(regions_df_monthly$months, levels = unique(regions_df_monthly$months))

regions_df_monthly %>% 
  ggplot(aes(x=months, group=1)) + 
  geom_bar(aes(y=avg_PPT_value/10, col = "Percipitation  [mm]"), stat = "identity", fill = "lightblue") +
  geom_line(aes(y=avg_temp_value, col="Avg. Temp [C]"), size = 1) + 
  geom_line(aes(y=avg_Tmax_value, col = "Max. Avg. Temp [C]")) +
  geom_line(aes(y=avg_Tmin_value, col = "Min. Avg.  Temp [C]")) +
  geom_point(aes(y=avg_temp_value)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust=0.5)) +
  facet_wrap(.~ecoRegion) +
  scale_color_manual(values=c("darkorange", "red" ,"blue", "black")) +
  scale_y_continuous(sec.axis = sec_axis(~.*10, name = "Percipitation [mm]")) +
  labs(x="month", y="Temperature [C]")

```
```{r}
# Perhaps we can make this look a bit nicer

ggplot(data = regions_df_monthly, mapping = aes(x = months, y = avg_temp, group = 1)) + 
  geom_bar(stat = "identity", color="blue", fill="blue", width = 0.5) + 
  geom_line(mapping = aes(y = celsius+30), color="red", size=1.5) + # Scale data to match desired scale
  scale_y_continuous("Precipitation [mm]", 
                     sec.axis = sec_axis(~ . -30, name = "Temperature [째C]") # Reverse transformation to match data
  )

# Looks like we need to 'normalize' temperature in order to work with the weird limits imposed by having two axes

# Transformation: Found on https://stackoverflow.com/questions/58774705/y-limits-for-ggplot-with-sec-axis
#y <- a + (Temp - mean(Temp)/sd(Temp)) * s

ylim.prim <- c(min(regions_df_monthly$avg_PPT_value), max(regions_df_monthly$avg_PPT_value))   # in this example, precipitation
ylim.sec <- c(min(regions_df_monthly$avg_Tmin_value), max(regions_df_monthly$avg_Tmax_value))    # in this example, temperature

b <- diff(ylim.prim)/diff(ylim.sec)
a <- ylim.prim[1] - b*ylim.sec[1] # there was a bug here

ggplot(regions_df_monthly, aes(months, avg_PPT_value)) +
  geom_col() +
  geom_line(aes(y = a + avg_temp_value*b), color = "red") +
  scale_y_continuous("Precipitation", sec.axis = sec_axis(~ (. - a)/b, name = "Temperature")) +
  facet_wrap(~ecoRegion) +
  scale_x_discrete("Month", breaks = 1:12) 
ggplot(regions_df_monthly, aes(months, avg_PPT_value)) +
  geom_line() + 
  geom_line(aes(y = (a + ((avg_temp_value - mean(TEMP))/sd(TEMP)) * s) ), color = "red") +
  scale_y_continuous("Precipitation", 
                     limits=ylim.prim,
                     sec.axis = sec_axis(~ (. - a) / s * sd(TEMP) + mean(TEMP), name = "Temperature"),) +
  scale_x_continuous("Month", breaks = 1:12) +
  theme(axis.title.y.right = element_text(colour = "red"))
```

```{r}
# Seasonal Version
regions_df_seasons <- regions_df %>% 
  filter(months %in% c("Winter", "Spring", "Summer", "Fall"))

regions_df_seasons$months <- as.character(regions_df_seasons$months)

regions_df_seasons$months <- factor(regions_df_seasons$months, levels = unique(regions_df_seasons$months))

regions_df_seasons %>% 
  ggplot(aes(x=months, group=1)) + 
  geom_bar(aes(y=avg_PPT_value/10, col = "Percipitation  [mm]"), stat = "identity", fill = "lightblue") +
  geom_line(aes(y=avg_temp_value, col="Avg. Temp [C]"), size = 1) + 
  geom_line(aes(y=avg_Tmax_value, col = "Max. Avg. Temp [C]")) +
  geom_line(aes(y=avg_Tmin_value, col = "Min. Avg.  Temp [C]")) +
  geom_point(aes(y=avg_temp_value)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust=0.5)) +
  facet_wrap(.~ecoRegion) +
  scale_color_manual(values=c("darkorange", "red" ,"blue", "black")) +
  scale_y_continuous(sec.axis = sec_axis(~.*10, name = "Percipitation [mm]")) +
  labs(x="month", y="Temperature [C]")


```

