---
title: "TNRS_resolution"
author: "JT Miller"
date: "2022-08-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### A markdown for exploring the package TNRS for plant taxonomic resolution 

Load Necessary Libraries
```{r}
library(TNRS)
library(tidyverse)
```

Bring in idigbio occurrence data for the western North America 
```{r}
idigbio_plants <- read.csv("/home/jt-miller/Globi-Bees-Plant-Interactions/Globi-Bees-Plants-Interactions/Raw-Data/Plant-Data/idigbio-plant-unresolved-names.csv")
```
Lets take a look at just plant species 
```{r}
idigbio_plants_unique_SN <- idigbio_plants %>% 
  distinct(dwc.scientificName, .keep_all = TRUE)
```


A problem with using the TNRS API is that we are only capable of resolving 5000 scientificNames during a query...considering we have 51,729 observations we need to run a for loop 11 times. Maybe not too bad? 

Getting to understand the TNRS functionality 
```{r}
# Comparing subsets
subset_100 <- idigbio_plants_unique_SN[1:100,10] # column 10 is the scientificName 
subset_1000 <- idigbio_plants_unique_SN[1:1000,10]
subset_10000 <- idigbio_plants_unique_SN[1:10000,10]
subset_5000 <- idigbio_plants_unique_SN[1:5000, 10]

tnrs_sources <- TNRS_sources() 

print(tnrs_sources) # Possible resources we may use. 

resolved_subset <- TNRS(subset_100, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )

# What happens when we provide multiple sources?
resolved_subset2 <- TNRS(subset_100, # data
                        sources = c("wfo", "tropicos"),# the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
# Maintains the same number of observations, however now some of the sources have been replaced with tropicos
resolved_subset2 %>% 
  group_by(Source) %>% 
  count()

tropicos_pulls <- resolved_subset2 %>% 
  filter(Source == "tropicos")

name_comp <- tropicos_pulls$Name_submitted

# Lets ask, why did it replace those records?

resolved_subset %>% 
  filter(Name_submitted %in% name_comp) %>% 
  mutate(Overall_score = as.integer(Overall_score)) %>% 
  summarise(Overall_score_sum = sum(Overall_score))

tropicos_pulls %>% 
   filter(Name_submitted %in% name_comp) %>% 
  mutate(Overall_score = as.integer(Overall_score)) %>% 
  summarise(Overall_score_sum = sum(Overall_score))

# Suggesting that the overall score improves by the inclusion of the tropicos source. 

resolved_subset %>% 
  filter(Name_submitted %in% name_comp)

tropicos_pulls

# We can also look at a small dataset such as this pretty easily and manually compare them. It does indeed increase the scores for adding tropicos as a taxonomic resource. 

# It seems probable that we could achieve the same results by changing the matches argument

resolved_subset3 <- TNRS(subset_100, # data
                        sources = c("wfo","tropicos"), # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "all", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
resolved_subset3 %>% 
  filter(Name_submitted %in% name_comp) %>% 
  select(Name_submitted, Overall_score, Source)

# Doesnt really seem like it...odd. As an example epilobium hallianum 

resolved_subset %>% 
  filter(Name_submitted == "epilobium hallianum")

resolved_subset %>% 
  filter()
resolved_subset2 %>% 
  filter(Name_submitted == "epilobium hallianum")

resolved_subset3 %>% 
  filter(Name_submitted == "epilobium hallianum") 

# Now lets try to understand what exactly the authors mean by the classification argument


resolved_subset4 <- TNRS(subset_100, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "tropicos", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "all", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )

resolved_subset5 <- TNRS(subset_100, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "all", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )


# Lets check if these two agree with eachother. 
classification_comparison <- ifelse(resolved_subset4$Accepted_species == resolved_subset5$Accepted_species, "Agree",
                                    ifelse(resolved_subset4$Accepted_species != resolved_subset5$Accepted_species, 
                                           "Disagree"))

unique(classification_comparison) # Seems that it is indeed the same whether using wfo or tropicos as the classification method. Which agrees with the documentation on the classification argument.
  
# Seems that classification can only be one argument maximum.

# 

# Warnings: https://tnrs.biendata.org/warnings/ Suggests that there are Partial Matches (1), meaning that Name matched is a higher taxon than the name submitted. 


# side note: If u feed the TNRS function a list of redundant names, itll return a list of distinct back. 
```

Next, Lets test where we will hit bottle necks with the API
```{r}
# the github issues page mentions that 1,000 was a soft cap made for the r version of the API, however it was updated to 5,000 at some point. Lets test this 
start.time <- Sys.time()
thousand_resolved <- TNRS(subset_1000, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken #33 seconds to resolve


start.time <- Sys.time()

fiveThousand_resolved <- TNRS(subset_5000, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
end.time<- Sys.time()

difference <- difftime(end.time, start.time, units = "minutes")
difference #150 seconds to resolve, appears that the resolution time is a linear function at least from there. 

# Now the big question, does it fail when trying to query more than 5000 records? 

subset_5001 <- idigbio_plants_unique_SN[1:5001,10] # column 10 is the scientificName

start.time <- Sys.time()
hardCap_resolved <- TNRS(subset_5001, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # 156 seconds to resolve, appears that the resolution time is a linear function at least from there. 


# And Test the limits further
subset_10000 <- idigbio_plants_unique_SN[1:10000,10] # column 10 is the scientificName

start.time <- Sys.time()
tenThousand_resolved <- TNRS(subset_10000, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # 307 seconds to resolve, appears that the resolution time is a linear function at least from there. 

# So it does work however you do generate a warning, not sure if there is any disagreement? I suppose we can check via manual

subset_5000to10000 <- idigbio_plants_unique_SN[5000:10000,10]

tenThousand2ndHalf_resolved <- TNRS(subset_5000to10000, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )

tenThousand1stHalf_resolved <- tenThousand_resolved[5000:10000,]

CapComparison <- ifelse(tenThousand1stHalf_resolved$Accepted_species == tenThousand2ndHalf_resolved$Accepted_species, "Agree",
                                    ifelse(tenThousand1stHalf_resolved$Accepted_species != tenThousand2ndHalf_resolved$Accepted_species, 
                                           "Disagree"))

unique(CapComparison) # All comparisons are equivalent. 
```
For interests sake, can we write a for-loop to look at time it takes to resolve names? 
```{r}
x <- rep(NA, 10) # Create a holding vector to store our time 
yset <- c(1000, 2000, 3000, 4000, 5000, 6000, 7000, 8000, 9000, 10000) # Y will be the number of names pulled each query

yset <- as.integer(yset)
for(i in 1:length(yset)){
  
  y <- yset[i]
  for(j in 1:length(x)){
plant_name_subset <- idigbio_plants_unique_SN[1:max(y),10]
  
start.time <- Sys.time()
resolved_plants <- TNRS(plant_name_subset, # data
                        sources = "wfo", # the sources of the TNRS sources you wish to use for resolution. 
                        classification = "wfo", # Family classification to use...honestly I dont understand what they mean
                        mode = "resolve", # Options are to either "resolve" or "parse" the scientificName
                        matches = "best", # Options are to retunr "all" or "best" 
                        accuracy = NULL # Numeric: If specified you can set a value x <= score for what is returned. 
                        )
end.time <- Sys.time()
difference <- difftime(end.time, start.time, units = "secs")
x[j] <- difference #33 seconds to resolve
}

return(list(x))
  
}

plot(yset, x)






```

