---
title: "LCVP_resolution"
author: "JT Miller"
date: "2022-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### A markdown to look at taxonomic resolution using the Leipzig Catalogue of Vascular Plants
https://www.nature.com/articles/s41597-020-00702-z.pdf 
https://github.com/idiv-biodiversity/lcvplants

Libraries
```{r}
# Required Librariers
library(devtools)
#devtools::install_github("idiv-biodiversity/LCVP") # Apparently this database is large enough to error R out if you have default download options. Set options(timeout=9999999) to remedy this issue and avoiding a FAIL errorcode. 

#devtools::install_github("idiv-biodiversity/lcvplants")
# Load the package
library(lcvplants)
library(tidyverse)
```

Run an example to make sure that the package works 
```{r}
# single names
lcvp_search("Hibiscus vitifolius") # Note the warning says the first accepted name was returned. 
lcvp_fuzzy_search("Hibiscus vitifolius") # Same as lcvp_search, but returns ALL matches including synonyms 

# Can also include infra specific names and authorities
lcvp_search("Hibiscus abelmoschus var. betulifolius Mast.")

# Also works on vectors of names
search_result <- lcvp_search(c("Hibiscus abelmoschus var. betulifolius Mast.", "Hibiscus abutiloides Willd.", 
       "Hibiscus aculeatus", "Hibiscus acuminatus"))

# You can see the summary results
lcvp_summary(search_result)

```
The package can also detect misspellings by specifying a max_distance argument. 
```{r}
# set max_distance = 2 meaning there can be at most 2 letters misspelled in the genus, then epithet 
lcvp_search("Hibiscus vitifolia", max_distance = 2) # Note vitifolia vs vitifolius

# It appears that the specific epithet can be pretty misspelled, however the any genus misspellings will void the search
lcvp_search("Hibiscus viticooiua", max_distance = 3)
lcvp_search("Libiscus vitifolius", max_distance = 1) # This is different then what the authors specified in their methods, possibly something changed? 
```
Searches can be also made based upon Order, Family, Genus, or Author.
```{r}
lcvp_group_search("AA", search_by = "Genus") # Look for genus AA 

lcvp_group_search(c("Orchidaceae", "Poaceae", "Phyllanthaceae"), search_by = "Family") # Look up the following families
```
The package also allows for comparisons to be made between two lists of species
```{r}
# Make a sp list 1
splist1 <- sample(LCVP::tab_lcvp$Input.Taxon[1:10]) # Take the first 10 rows of the tab_lcvp df and subset it into a vector of the first 10 values in Input.Taxon
# Make a sp list 2
splist2 <- sample(LCVP::tab_lcvp$Input.Taxon[1:10]) # ""

lcvp_match(splist1, splist2)



```
Finally, this package allows us to join two tables based upon vascular plant names
```{r}
splist1 <- sample(LCVP::tab_lcvp$Input.Taxon[2:10]) # Take the 2:10 rows of the tab_lcvp df and subset it into a vector of the first 2:10 values in Input.Taxon (9 values)
x <- data.frame("Species" = splist1, "Trait1" = runif(length(splist1))) # Create a dummy dataframe with the a column for Species made with the vector splist1, then populate a second column with uniform distribution of numbers for the length of the vector splist1

splist2 <- sample(LCVP::tab_lcvp$Input.Taxon[1:8]) # Grab the first 8 rows
y <- data.frame("Species" = splist2,  "Trait2" = runif(length(splist2)), "Trait3" = runif(length(splist2))) # Add in the compoinent of a third trait

# a Full Join with the two tables
lcvp_join(x, y, c("Species", "Species"), type = "full") # Note that the comparison generates NAs for where they dont match up. For example list 2 does not contain the elements 2 & 6 found in list 1. 

# A Left join (Only return values in x)
lcvp_join(x, y, c("Species", "Species"), type = "left") # Should give us 9 values

# A Right join (Only return values in y)
lcvp_join(x,y, c("Species", "Species"), type = "right") # Should give us 8 values

# A Inner join (Only return values that exist in both x & y)
lcvp_join(x,y, c("Species", "Species"), type = "inner") # Should give us 7 values
```
### Enough test driving, lets try this out on some real data

```{r}
idigbio_plants <- read.csv("/home/jt-miller/Globi-Bees-Plant-Interactions/Globi-Bees-Plants-Interactions/Raw-Data/Plant-Data/idigbio-plant-unresolved-names.csv")

idigbio_plants2 <- read.csv("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/iDigBio-Pull/Unresolved-iDigBio-Plants.csv")

idigbio_plants_unique_SN <- idigbio_plants %>% 
  distinct(dwc.scientificName, .keep_all = TRUE)

idigbio_plants_unique_SN %>% 
  filter(dwc.occurrenceID == "c7fc7dec-ece0-4344-b351-7f75319ad3a1")

# Create some subsets to test query time
subset_100 <- idigbio_plants_unique_SN[1:100,10] # column 10 is the scientificName 
subset_1000 <- idigbio_plants_unique_SN[1:1000,10]
subset_10000 <- idigbio_plants_unique_SN[1:10000,10]
subset_5000 <- idigbio_plants_unique_SN[1:5000, 10]
```

```{r}
search_result <- lcvp_search(subset_100) # Error occurred suggesting binomial name violation, check for example "arctostaphylos" 
subset_100_a <- lapply(subset_100, function(z) z[grepl("arctostaphylos", z)])
subset_100_a # Issue with element 6, no epithet specified. Lets write some code that with remove any entries that don't have more than a a singular word present 

test_df <- idigbio_plants_unique_SN[sapply(strsplit(as.character(idigbio_plants_unique_SN$dwc.scientificName)," "),length)>1,] # Drops anything with less than 2 words in the character string  

test_df2 <- idigbio_plants_unique_SN[sapply(strsplit(as.character(idigbio_plants_unique_SN$dwc.scientificName)," "),length)==1,] # Shows the ones we got rid of, besides that one value that has zero

test_df3 <- idigbio_plants_unique_SN[sapply(strsplit(as.character(idigbio_plants_unique_SN$dwc.scientificName)," "),length)==0,]

test_df <- test_df[1:100, 10]
test_df <- lapply(test_df,  function(z) z[grepl("arctostaphylos", z)])

test_df # Seems to have fixed the issue 


### Apply this to our actual DF 
idigbio_plants_unique_SN_db <- idigbio_plants_unique_SN[sapply(strsplit(as.character(idigbio_plants_unique_SN$dwc.scientificName)," "),length)>1,]

### Its also important to keep these genus level observations for later in case we need them 
idigbio_plants_genus_lvl_obs <- idigbio_plants_unique_SN[sapply(strsplit(as.character(idigbio_plants_unique_SN$dwc.scientificName)," "),length)==1,]

### And grab 2 problematic occurrences with '.' that are only genus level observations and add them to this list
problematic_genus_records <-idigbio_plants_unique_SN_db %>%
  filter(dwc.scientificName == "cf. ricinus" |  dwc.scientificName == "aff. alnus")
# cf. ricinus Is not a valid name and is having issues, not sure outside of just dropping it. https://www.idigbio.org/portal/records/e8c3a500-647b-446a-b372-b17c11592f3d
# aff. alnus is also an issue, is it because of the special character '.' in the genus?
### rbind them

idigbio_plants_genus_lvl_obs <- rbind(idigbio_plants_genus_lvl_obs, problematic_genus_records)

### Get rid of the ones in our dbl dataset
idigbio_plants_unique_SN_db <- idigbio_plants_unique_SN_db %>% 
  filter(!(dwc.scientificName == "cf. ricinus" |  dwc.scientificName == "aff. alnus"))

idigbio_plants_unique_SN_db %>% 
  filter(grepl("cf\\. ricinus", dwc.scientificName))
```
Write our genus level obs to a .csv and store them for later in case their needed 
```{r eval=FALSE, include=FALSE}
write.csv(idigbio_plants_genus_lvl_obs, "/home/jt-miller/Globi-Bees-Plant-Interactions/Processed-Data/idigbio-plants-genus-lvl-obs-only.csv", row.names = FALSE)
```

```{r}
# Subsetting our new df 
subset_100 <- idigbio_plants_unique_SN_db[1:100,10] # column 10 is the scientificName 
subset_1000 <- idigbio_plants_unique_SN_db[1:1000,10]
subset_10000 <- idigbio_plants_unique_SN_db[1:10000,10]
subset_5000 <- idigbio_plants_unique_SN_db[1:5000, 10]

start.time <- Sys.time()
search_result <- lcvp_search(subset_100) # Alright, successful output. 
# Note notation: PL_Comparison = Plant List Comparison
# PL_Alternative = provides the alternative name from the The Plant List 
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # 1.5 seconds to resolve 100 names 


start.time <- Sys.time()
search_result <- lcvp_search(subset_1000) # Alright, successful output. 
# Note notation: PL_Comparison = Plant List Comparison
# PL_Alternative = provides the alternative name from the The Plant List 
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # 18 seconds to resolve

start.time <- Sys.time()
search_result <- lcvp_search(subset_10000) # Alright, successful output. 
# Note notation: PL_Comparison = Plant List Comparison
# PL_Alternative = provides the alternative name from the The Plant List 
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # 4.78 minutes to resolve

# Trying our full dataset
start.time <- Sys.time()
LCVP_resolved_idigbio_plants <- lcvp_search(idigbio_plants_unique_SN_db$dwc.scientificName) # Failed to run due to taxonomic name issue
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken
```
Another Taxonomic Name issue 
```{r}
idigbio_plants_unique_SN_db %>%
  filter(dwc.scientificName == "achnatherum x") # Taxon Match Fail 
#  achnatherum x full name should be achnatherum x bloomeri (Bolander) Barkworth, note that this wasn't the parsing script that did this, it was like this from the pull. https://www.idigbio.org/portal/records/2c012181-5b72-463e-b484-8f2f7cf12c37


idigbio_plants_unique_SN_db_fixed <- idigbio_plants_unique_SN_db %>% 
  mutate(dwc.scientificName = replace(dwc.scientificName, dwc.scientificName ==  "achnatherum x" ,"achnatherum x bloomeri (Bolander) Barkworth" ))

idigbio_plants_unique_SN_db_fixed %>%
  filter(dwc.scientificName == "achnatherum x") # Doesnt appear

idigbio_plants_unique_SN_db_fixed %>%
  filter(dwc.scientificName == "achnatherum x bloomeri (Bolander) Barkworth") # Confirmed Fix


```
```{r}
# Retry full dataset 
start.time <- Sys.time()
LCVP_resolved_idigbio_plants <- lcvp_search(idigbio_plants_unique_SN_db_fixed$dwc.scientificName) # Failed to run due to taxonomic name issue
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # Something errors out somewhere down the line...odd the error is very general as well so Not super helpful for debugging
```
Proceeding to cut the df into 6 sections in order to debug 
```{r}
splitter<-function(x,groups=y){
  dd<-as.data.frame(x)
  dd$split<-sample(groups,size=nrow(dd),replace=T)
  for(i in 1:groups){
    ddd<-dd[dd$split==i,]
    assign(paste0("Plants_for_Resolution",i),ddd,envir=globalenv())
  }
}
splitter(idigbio_plants_unique_SN_db_fixed, groups = 6)

start.time <- Sys.time()
LCVP_resolved_idigbio_plants1 <- lcvp_search(Plants_for_Resolution1$dwc.scientificName)
LCVP_resolved_idigbio_plants2 <- lcvp_search(Plants_for_Resolution2$dwc.scientificName)
LCVP_resolved_idigbio_plants3 <- lcvp_search(Plants_for_Resolution3$dwc.scientificName)
LCVP_resolved_idigbio_plants4 <- lcvp_search(Plants_for_Resolution4$dwc.scientificName)
LCVP_resolved_idigbio_plants5 <- lcvp_search(Plants_for_Resolution5$dwc.scientificName)
LCVP_resolved_idigbio_plants6 <- lcvp_search(Plants_for_Resolution6$dwc.scientificName)
end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken
```
Issues with the 5th subset, try using unique to see if theres a simplistic solution
```{r eval=FALSE, include=FALSE}
unique(Plants_for_Resolution5$dwc.scientificName) # Thats a lot of results and nothing immediately noticable, lets try splitting further to reduce the problem to seeable chunks

# Modifying our splitter function a tad for more broad naming usage 
splitter<-function(x,groups=y, output_name){
  dd<-as.data.frame(x)
  dd$split<-sample(groups,size=nrow(dd),replace=T)
  for(i in 1:groups){
    ddd<-dd[dd$split==i,]
    assign(paste0(output_name,i),ddd,envir=globalenv())
  }
}

splitter(Plants_for_Resolution5, groups = 2, output_name = "subset_5_split")

LCVP_resolved_idigbio_plants5_split1 <- lcvp_search(subset_5_split1$dwc.scientificName) # Appears to have failed on the first half, checking the second half just to make sure the errors dont exist there as well. 
LCVP_resolved_idigbio_plants5_split2 <- lcvp_search(subset_5_split2$dwc.scientificName) # Appears both contain issues...interesting 


# Split futher to isolate the issue?

splitter(subset_5_split1, groups = 5, output_name = "subset_5_splitfurther_1.") # Split further into sets of ~1000
LCVP_resolved_idigbio_plants5_split1.1 <- lcvp_search(subset_5_splitfurther_1.1$dwc.scientificName) # Fails
LCVP_resolved_idigbio_plants5_split1.2 <- lcvp_search(subset_5_splitfurther_1.2$dwc.scientificName) # Succeeds
LCVP_resolved_idigbio_plants5_split1.3 <- lcvp_search(subset_5_splitfurther_1.3$dwc.scientificName) # Succeeds
LCVP_resolved_idigbio_plants5_split1.4 <- lcvp_search(subset_5_splitfurther_1.4$dwc.scientificName) # Succeeds
LCVP_resolved_idigbio_plants5_split1.5 <- lcvp_search(subset_5_splitfurther_1.5$dwc.scientificName) # Succeeds

# Something is causing the error out in the first ~850 names, try unique just for scanning purposes 
print(subset_5_splitfurther_1.1$dwc.scientificName)

# Nothing immediately obvious, I recall in their paper they mentioned that the function couldn't operate when there was special characters. 

questionable_characters <- subset(subset_5_splitfurther_1.1, grepl("\\?", subset_5_splitfurther_1.1$dwc.scientificName)) # Seems there is a problematic record where a user inputted a '?' and a '&' for that matter. Lets see if we can just remove this record for now 

subset_5_splitfurther_1.1_d <- subset_5_splitfurther_1.1 %>% 
  filter(!grepl("\\?", dwc.scientificName))

# Check if it runs now 
LCVP_resolved_idigbio_plants5_split1.1 <- lcvp_search(subset_5_splitfurther_1.1_d$dwc.scientificName) # Still fails, another issue is present


# Try out looking for commas? 
subset_5_splitfurther_1.1_d <- subset_5_splitfurther_1.1 %>% 
  filter(grepl("\\,", dwc.scientificName)) # Try their removals 
# Possibly: 3fa37b3a-2679-4563-906c-e4aa1f885d10 [not], 
# 4f31df99-6fba-4213-9568-8ad403543a88 [YES] , 	
# cb5c80b2-ea6f-4e80-8da8-3550b2e24f62 [not]

subset_5_splitfurther_1.1_d <- subset_5_splitfurther_1.1 %>% 
  filter(!grepl("\\,", dwc.scientificName))

# Another test 
subset_5_splitfurther_1.1_d <- subset_5_splitfurther_1.1 %>% 
  filter(coreid == "4f31df99-6fba-4213-9568-8ad403543a88") 


print(subset_5_splitfurther_1.1_d$dwc.scientificName) # Not sure what the issue is here, var.? 

subset_5_splitfurther_1.1_d <- subset_5_splitfurther_1.1 %>% 
  filter(grepl("var\\.", dwc.scientificName)) # NOT var.  ... but maybe its because it has 2 instances of var. ? 


LCVP_resolved_idigbio_plants5_split1.1 <- lcvp_search(subset_5_splitfurther_1.1_d$dwc.scientificName) # It seems issues exist when x is involved? Can we see if that is true? 

# Lets see if we can construct a grepl argument to retrieve records where there is a ' x ' since that seems to be an issue?

# Some cool code I want to keep, thought it wasn't helpful in this circumstance. 

# subset_5_splitfurther_1.1_d <- subset_5_splitfurther_1.1 %>% 
 #  filter(grepl(" x +[a-z]\\.", dwc.scientificName)) 

### Comparison to other df that failed ###

subset_5_split2_d <- subset_5_split2 %>% 
  filter(grepl("\\,", dwc.scientificName))

subset_5_split2_d <- subset_5_split2 %>% 
  filter(coreid == "4f31df99-6fba-4213-9568-8ad403543a88") # Same error out! 

subset_5_split2_d <- subset_5_split2_d %>% 
  filter(!coreid == "4f31df99-6fba-4213-9568-8ad403543a88") # Same error out! 


LCVP_tests2 <- lcvp_search(subset_5_split2_d$dwc.scientificName) # Aha! its this one really annoying record. Possibly due to .var showing up twice. 

```
Applying the fix to subset 5
```{r}
# Find the .var doubles
Plants_for_Resolution5_d <- Plants_for_Resolution5 %>% 
  filter(grepl("\\,", dwc.scientificName)) %>% 
  filter(grepl(".var", dwc.scientificName))

# Remove 

Plants_for_Resolution5_d <- Plants_for_Resolution5 %>% 
  filter(!coreid == "44f31df99-6fba-4213-9568-8ad403543a88")

Plants_for_Resolution5_d  <- Plants_for_Resolution5_d %>% 
  filter(!grepl("cf. ricinus", dwc.scientificName)) # Not sure how this record found its way back...? Possibly coding error? 

# Run the LCVP search

LCVP_resolved_idigbio_plants5 <- lcvp_search(Plants_for_Resolution5_d$dwc.scientificName)
```
Apply the method to fix subset 6
```{r}
Plants_for_Resolution6_d <- Plants_for_Resolution6 %>% 
  filter(grepl("\\?", dwc.scientificName))

# Unsuccessful, attempt the splitting method to isolate it?
splitter(Plants_for_Resolution6, groups = 10, output_name = "subset_6_split1.")

LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.1$dwc.scientificName) # Succeeded 
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.2$dwc.scientificName) # Succeeded
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.3$dwc.scientificName) # Succeeded 
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.4$dwc.scientificName) # Failed
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.5$dwc.scientificName) # Succeeded
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.6$dwc.scientificName) # Succeeded
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.7$dwc.scientificName) # Succeeded
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.8$dwc.scientificName) # Succeded 
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.9$dwc.scientificName) # Succeeded
LCVP_resolved_idigbio_plants6 <-lcvp_search(subset_6_split1.10$dwc.scientificName) # Succeeded 


# Data we know has an issue 
subset_5_split2_d <- subset_5_split2 %>% 
  filter(grepl("\\,", dwc.scientificName)) # Tester 

# The function for testing resolution with lcvp
LCVP_res_tester <- function(x){
  resolved <- lcvp_search(x)
}
# Our test dataset
test_list <- list(subset_5_split2_d$dwc.scientificName)

### The catcher ###
flag <- integer()
start.time <- Sys.time()
 
for(i in 1:lengths(test_list)){
  tryCatch({
   # print(i)
    LCVP_res_tester(x = test_list[[1]][i])
  }, error = function(e){cat("Error T/F statement :", conditionMessage(e), "\n")
    flag <<- c(flag,i)
    }
  )
}
print(flag)

end.time<- Sys.time()
time.taken <- end.time - start.time
time.taken # 4.78 minutes to resolve
```


```{r}
### The Catcher function ###
LCVP_catcher <- function(subset_list){
  
  LCVP_res_tester <- function(x){ # Make our tester function
  resolved <- lcvp_search(x)
  }
  
  flag <- integer() # Create list numeric labeling for identifying which SN are erroring 
  
  # Create a for loop that uses tryCatch to find errors, runs the subset fully to the end however. 
  for(i in 1:lengths(subset_list)){
  tryCatch({
   # print(i)
    LCVP_res_tester(x = subset_list[[1]][i])
  }, error = function(e){cat("Error T/F statement :", conditionMessage(e), "\n")
    flag <<- c(flag,i)
    }
  )
}
print(flag)
  
}
```



```{r warning=FALSE}
### Using The catcher on our *Problematic* subsets (note that the whole dataset would take way too long ~60 hours computation time) since we're fancy ### Note that the warnings are silenced in this chunk

subset_6_name_list <- list(subset_6_split1.4$dwc.scientificName)

### The catcher ###
flag <- integer()

for(i in 1:lengths(subset_6_name_list)){
  tryCatch({
   # print(i)
    LCVP_res_tester(x = subset_6_name_list[[1]][i])
  }, error = function(e){cat("Error T/F statement :", conditionMessage(e), "\n")
    flag <<- c(flag,i)
    }
  )
}
print(flag)

```

```{r}
# Memory requirements
object.size(idigbio_plants_unique_SN_db_fixed)/2^20 # MB conversion 

```
Parallelize to cut down on computation time
```{r}
library(doParallel)
```

```{r}
# Example of how we use doParallel
numCores <- detectCores()
registerDoParallel(numCores)

foreach(i = 1:3) %dopar% {
  sqrt(i)
}
```

```{r}
# Trying it on our data now 
subset_6_name_list <- list(subset_6_split1.4$dwc.scientificName)

### The catcher ###
flag <- integer()

system.time({
foreach(for(i in 1:lengths(subset_6_name_list)){
  tryCatch({
   # print(i)
    LCVP_res_tester(x = subset_6_name_list[[1]][i])
  }, error = function(e){cat("Error T/F statement :", conditionMessage(e), "\n")
    flag <<- c(flag,i)
    }
  )
})
print(flag)
})


subset_list <- list(subset_6_split1.4$dwc.scientificName)
subset_list <- list("clematis occidentalis var. grosseserrata (rydb.) j.s.pringle", "castilleja wightii", "triglochin concinna var. debilis (m.e. jones) j.t. howell, orth. var.")
flag <- integer()
foreach(i = 1:lengths(subset_list)) %dopar% {
  tryCatch({
   # print(i)
    LCVP_res_tester(x = subset_list[[1]][i])
  }, error = function(e){cat("Error T/F statement :", conditionMessage(e), "\n")
    flag <<- c(flag,subset_list[[1]][i])
    }
  )
  print(flag)
 
}

test <- lcvp_search("triglochin concinna var. debilis (m.e. jones) j.t. howell, orth. var.")


```


```{r}

subset_list <- list("clematis occidentalis var. grosseserrata (rydb.) j.s.pringle", "castilleja wightii", "triglochin concinna var. debilis (m.e. jones) j.t. howell, orth. var.", "clematis occidentalis var. grosseserrata (rydb.) j.s.pringle", "clematis occidentalis var. grosseserrata (rydb.) j.s.pringle", "clematis occidentalis var. grosseserrata (rydb.) j.s.pringle", "triglochin concinna var. debilis (m.e. jones) j.t. howell, orth. var.")
LCVP_res_tester <- function(x){
  resolved <- lcvp_search(x)
}

flag <- integer()
for(i in 1:length(subset_list)){
  tryCatch({
   # print(i)
    LCVP_res_tester(x = subset_list[[i]])
  }, error = function(e){cat("Error T/F statement:", conditionMessage(e), "\n", subset_list[[i]], "\n")
    #flag <<- c(flag,i)
    }
  )
}
print(flag)


idigbio_plants %>% 
  filter(coreid == "4f31df99-6fba-4213-9568-8ad403543a88")

```

