---
title: "Resolution_Hier"
author: "JT Miller"
date: "2022-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
opts_knit$set(root.dir = "/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution")

```

## A markdown to build a for-loop for resolution at a hierarchical level. 

Libraries 
```{r}
library(tidyverse)
library(taxize)
library(sqldf)
```
Bring in the data
```{r}
plant_names <- read.delim("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/distinct_c_names.txt", header = TRUE, quote = "")
```

Run it through GBIF's parsing service 
```{r}
gbif_parsed_plants <- gbif_parse(plant_names$scientificName) # Uses GBIF's API to parse names

# Re append Authorship where written out separately. Create an if-else statement, if ScientificNameAuthorship was not filled out by the providers, then look at the authorship column that was broken out by gbif-parse. If there is a author in plant_names$ScientificNameAuthorship prioritize this, if not use gbif_parsed_plants$authorship. No authorship present will result in NA.

gbif_parsed_plants$authorship <- ifelse(plant_names$scientificNameAuthorship != "", 
                                          plant_names$scientificNameAuthorship, gbif_parsed_plants$authorship)


```

Create a for-loop that calls upon Nomer bash script for resolution. Here, we want to design a hierarchical resolution tool that starts at the most detailed version of the name and takes a step back upon failure to resolve a name. 

Proposed Hierarchy: canonicalNameWithMarker + authorship > canonicalNameWithMarker > canonicalName > genusORAbove + specificEpithet

Basic Logic:

boolean found = false
if (authorship exists)
    found = try canonicalNameWithMarker + authorship
if (!found)
    found = try canonicalNameWithMarker
if (!found AND canonicalNameWithMarker != canonicalName)
    found = try canonicalName

generatedName = genusOrAbove + specificEpithet
if (!found AND canonicalName != generatedName)
    found = try genusOrAbove + specificEpithet
    
Note! Currently missing logic for dealing with names that will map to multiple resolutions given no authorship...Add this later!!!

```{r}
# First, lets use a small subset of names for resolution

name_subset <- gbif_parsed_plants # Grabbing ALL of the names


name_subset <- distinct(name_subset, scientificname, authorship, .keep_all = TRUE ) # Run this prior to using the shell script.

write.table(name_subset, "/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/name_subset.txt", row.names = FALSE, sep="\t", quote = FALSE)

nomer_w_authorship_cmd <- "/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/Nomer-w-authorship.sh"


### Defunct at the current moment ###

system2(nomer_w_authorship_cmd,  input = "name_subset1.txt")
# It appears this works OK, however; a secondary step could be to delete white space in authors names, this seems to be a commen reason for failure in matching. ex: M.E. Jones != M.E.Jones


# Remove whitespace...
name_subset2 <- name_subset1

name_subset2$authorship <- gsub("\\s+", '', name_subset2$authorship)

can_names <- name_subset$canonicalnamewithmarker
# Try a if-else logic...
exec_wait(nomer_w_authorship_cmd)

# Notes from 11/25: We need to make it where we paste in names from canonicalnamewithmarker & authorship into the shell cmd. 
if(!is.na(name_subset$authorship)){
  exec_wait(nomer_w_authorship_cmd)
}


# Trying the sys package
library(sys)

exec_wait(nomer_w_authorship_cmd)

_# Read what our output is giving us...


```


Table that plan. Seems that running bash through R has some limitations that I do not fully understand as of yet. Instead, for a short term solution, we can run Nomer on each type of iteration of the name, and then pull resolvedNames from the created tables in the hierarchical fashion mentioned. 

*Refer to the Nomer-loop.sh for running resolution*

First, document some of the changes necessary to build our iterations (Code chunk Currently Disabled)
```{r eval=FALSE, include=FALSE}
# Remove whitespace from subsetted data 
name_subset_wout_Awhitespace <- name_subset # create a copy
name_subset_wout_Awhitespace$authorship <- gsub("\\s+", '', name_subset$authorship) # remove whitespcae

# Write it out...
write.table(name_subset_wout_Awhitespace, file = "/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/name_subset-wout-Awhitespace.txt", row.names = FALSE, sep="\t", quote = FALSE)

```

Now lets load in each table in order of hierarchy. 
```{r}
# CanonicalNameWithMarker + Authorship Resolution
CNWM_A_Res <- read.delim("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/CNWM-A-resolution.txt", header = TRUE, quote = "")

# Change NONE to relationName
CNWM_A_Res <- CNWM_A_Res %>% 
  rename(relationName = NONE, resolvedName = canonicalnamewithmarker.1)

# CanonicalNameWithMarker + Authorship WhiteSpace removed Resolution
CNWM_AWS_Res <- read.delim("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/CNWM-AwoutWSP-resolution.txt", header = TRUE, quote = "")
CNWM_AWS_Res <- CNWM_AWS_Res %>% 
  rename(relationName = NONE, resolvedName = canonicalnamewithmarker.1)

# CanonicalNameWithMarker without Authorship Resolution
CNWM_only <-read.delim("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/CNWM-resolution.txt", header = TRUE, quote = "")
CNWM_only_Res <- CNWM_only %>% 
  rename(relationName = NONE, resolvedName = canonicalnamewithmarker.1)

# CanonicalName Resolution
CN_only <-read.delim("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/CN-resolution.txt", header = TRUE, quote = "")

CN_only_Res <- CN_only %>% 
  rename(relationName = NONE, resolvedName = canonicalname.1)


# Genus + specificEpithet resolution
SP_only <-read.delim("/home/jt-miller/Globi-Bees-Plant-Interactions/Plant-Data/Nomer-wfo-resolution/Final-Resolution/Hierarchy-Resolution/SP-resolution.txt", header = TRUE, quote = "")

SP_only_Res <- SP_only %>% 
  rename(relationName = NONE, resolvedName = ID.1)

```

Now build a Resolved Table, this table will pull names from the Nomer Resolutions.

Requirements:
Preference in resolution:
CanonicalNameWithMarker + Authorship > CanonicalNameWithMarker w/out Authorship > CanonicalName > Genus + specificEpithet
Capture the level at which resolution first occurs, then label it in a new field called resolutionType
Note when multiple resolved names exist for a resolution, label this in a boolean TRUE/FALSE field called multMaps
```{r eval=FALSE, include=FALSE}


CNWM_A_Res_d <- CNWM_A_Res %>% 
  distinct(canonicalnamewithmarker, authorship, relationName, canonicalnamewithmarker.1, .keep_all = TRUE)
# Construct a new column of resolved names appended to our name_subset_table, pull in order of prefence as listed above
name_resolved_table <- name_subset_table %>% 
  mutate(resolvedName = case_when(CNWM_A_Res_d$relationName != "NONE" ~ paste(CNWM_A_Res_d$canonicalnamewithmarker.1)
    
  ))


```
Using SQL Coalesce
```{r}
# Bring in base parsed table. Remove noninformative fields 
name_subset_table <- name_subset %>% 
  select(!14:22) %>%  # removes excess fields
  distinct(scientificname, type, canonicalnamecomplete, .keep_all = TRUE)

# First, change all of the NONE values to NA (NOMER syntax Correction)
CNWM_A_Res[CNWM_A_Res == "NONE"] <- NA
CNWM_AWS_Res[CNWM_AWS_Res == "NONE"] <- NA
CNWM_only_Res[CNWM_only_Res == "NONE"] <- NA
CN_only_Res[CN_only_Res == "NONE"] <- NA
SP_only_Res[SP_only_Res == "NONE"] <- NA
# By default NOMER still returns a the original verbatim name inputted even when resolution fails. Change this syntax via a new field called resolvedNameCorrected by using a CASE statement to make it clear these are NOT resolutions.

CNWM_A_Res <- sqldf("SELECT *,
                     CASE
                     WHEN relationName NOT NULL THEN resolvedName
                     WHEN relationName ISNULL THEN NULL
                     END AS resolvedNameCorrected
                     FROM CNWM_A_Res")

CNWM_AWS_Res <- sqldf("SELECT *,
                     CASE
                     WHEN relationName NOT NULL THEN resolvedName
                     WHEN relationName ISNULL THEN NULL
                     END AS resolvedNameCorrected
                     FROM CNWM_AWS_Res")

CNWM_only_Res <- sqldf("SELECT *,
                     CASE
                     WHEN relationName NOT NULL THEN resolvedName
                     WHEN relationName ISNULL THEN NULL
                     END AS resolvedNameCorrected
                     FROM CNWM_only_Res")

CN_only_Res <- sqldf("SELECT *,
                     CASE
                     WHEN relationName NOT NULL THEN resolvedName
                     WHEN relationName ISNULL THEN NULL
                     END AS resolvedNameCorrected
                     FROM CN_only_Res")

SP_only_Res <- sqldf("SELECT *,
                     CASE
                     WHEN relationName NOT NULL THEN resolvedName
                     WHEN relationName ISNULL THEN NULL
                     END AS resolvedNameCorrected
                     FROM SP_only_Res")

# Additional Mutation required for name_table_subset to match IDs with SP_only_Res: Combine genusOrAbove + specificEpithet 
name_subset_table <- name_subset_table %>% 
  unite('species', genusorabove:specificepithet, sep = " ")

# Change the name of SP_only_Res to match 
SP_only_Res <- SP_only_Res %>% 
  rename(species = ID)


# Now try using COALESCE METHOD in SQL 

# First Combine our Tables....
Combined_names <- sqldf("SELECT DISTINCT name_table.canonicalnamewithmarker AS canonicalNameWithMarker, first_res.relationName AS firstRelation, first_res.resolvedNameCorrected AS firstResolvedName, name_table.scientificname AS verbatimScientificName, second_res.relationName AS secondRelation, second_res.resolvedNameCorrected AS secondResolvedName, third_res.relationName AS thirdRelation, third_res.resolvedNameCorrected AS thirdResolvedName, fourth_res.relationName AS fourthRelation, fourth_res.resolvedNameCorrected AS fourthResolvedName, fifth_res.relationName AS fifthRelation, fifth_res.resolvedNameCorrected AS fifthResolvedName
                                           FROM name_subset_table AS name_table
                                           LEFT JOIN CNWM_A_Res AS first_res ON name_table.canonicalnamewithmarker = first_res.canonicalnamewithmarker
                                           LEFT JOIN CNWM_AWS_Res AS second_res ON name_table.canonicalnamewithmarker = second_res.canonicalnamewithmarker
                                           LEFT JOIN CNWM_only_Res AS third_res ON name_table.canonicalnamewithmarker = third_res.canonicalnamewithmarker
                                           LEFT JOIN CN_only_Res AS fourth_res ON name_table.canonicalname = fourth_res.canonicalname
                                           LEFT JOIN SP_only_Res AS fifth_res ON name_table.species = fifth_res.species")

# Use COALESCE to Create a relationFound Field
coalesced_names <- sqldf("SELECT verbatimScientificName, 
                                  canonicalNameWithMarker, firstRelation, secondRelation, thirdRelation, fourthRelation, fifthRelation, firstResolvedName, secondResolvedName, thirdResolvedName, fourthResolvedName, fifthResolvedName,
                          COALESCE(firstRelation, secondRelation, thirdRelation, fourthRelation, fifthRelation) relationFound
                         
                          FROM Combined_names")
# Create a CASE statement to identify at which level relation was found using COALESCE via a field called whenRelationFound
coalesced_org <- sqldf("SELECT *, 
                       CASE 
                       WHEN firstRelation NOT NULL THEN 'CNWMandAuthor'
                       WHEN secondRelation NOT NULL THEN 'CNWMandAuthorWSP'
                       WHEN thirdRelation NOT NULL THEN 'CNWMonly'
                       WHEN fourthRelation NOT NULL THEN 'CNonly'
                       WHEN fifthRelation NOT NULL THEN 'SPonly'
                       END AS whenRelationFound
                       FROM coalesced_names")

# Create a CASE statement to identify what the final resolved Name should be via a field called finalResolution 
coalesced_resolution <- sqldf("SELECT DISTINCT *, 
                              CASE 
                              WHEN whenRelationFound = 'CNWMandAuthor' THEN firstResolvedName
                              WHEN whenRelationFound = 'CNWMandAuthorWSP' THEN secondResolvedName
                              WHEN whenRelationFound = 'CNWMonly' THEN thirdResolvedName
                              WHEN whenRelationFound = 'CNonly' THEN fourthResolvedName
                              WHEN whenRelationFound = 'SPonly' THEN fifthResolvedName
                              END AS finalResolution
                              FROM coalesced_org")
coalesced_resolution <- distinct(coalesced_resolution, verbatimScientificName, canonicalNameWithMarker, finalResolution, .keep_all = TRUE)

# Now we need to identify where multiple mappings Occur, and whether resolution should even happen for these circumstances (i.e. is author needed?)
coalesced_resolution <- coalesced_resolution %>% 
  group_by(verbatimScientificName) %>% 
  mutate(multipleMappings = case_when(
    !n_distinct(finalResolution) == 1 ~ TRUE,
    TRUE ~ FALSE))


# Now look at when relation is found relative to coalescent resolution. If its found at the same time, this is uninformative and therefore 'bad' to keep
multmaps <- coalesced_resolution %>% 
  filter(multipleMappings == TRUE) %>% 
  group_by(verbatimScientificName) %>% 
  filter(!n_distinct(whenRelationFound) == 1) 
# This never occurs, and therefore is ambigious so we're better off discluding these names. 

# Test to see if multiple maps do indeed occur at the same whenRelationFound


```

```{r}
# Check to see what names were dropped? 



```

