---
title: "ClimatePullMarkdown"
author: "JT Miller"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A remake of my previous climateNA-Ecoregions markdown due to fatal issues associated with the markdown. 

Necessary Libraries
```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(zoom)
library(elevatr)
library(sp)
```

Load in the shapefiles 
```{r}
# Bring in the shape file for the North American Continent, assign it the WGS84 coordinate reference system
NA_eco_map <- sf::read_sf("/home/jt-miller/Documents/CCBER/occurrence-maps/filter_polygon/ecoregions-lvl3-NA/NA_CEC_Eco_Level3/NA_CEC_Eco_Level3.shp") %>% 
  sf::st_transform('+proj=longlat +datum=WGS84') 
```

Filter out the eco_map to only include the ecoregions of interest for this study (All of those that touch California)
```{r}
cali_ecoregions <- NA_eco_map %>% 
  filter(NA_L3NAME %in% c("Northern Basin and Range", "Central Basin and Range", "Mojave Basin and Range", "Sonoran Desert", "Baja California Desert", "California Coastal Sage, Chaparral, and Oak Woodlands", "Central California Valley", "Southern and Baja California Pine-Oak Mountains", "Sierra Nevada", "Klamath Mountains", "Coast Range", "Cascades", "Eastern Cascades Slopes and Foothills"))

# Make a california boundary shp

cali_boundary <- sf::read_sf("/home/jt-miller/Documents/CCBER/occurrence-maps/filter_polygon/Cali-Border/ca-state-boundary/CA_State_TIGER2016.shp") %>% 
  sf::st_transform('+proj=longlat +datum=WGS84')

# Make a color palette 
colpal_ext <- colorFactor(c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a", "#ffff99", "#b15928"), cali_ecoregions$NA_L3NAME)
```


# Use the sf package to create even spaced 4x4 (km) grid throughout our ecoregion shapefiles
```{r}
# Using https://stackoverflow.com/questions/53789313/creating-an-equal-distance-spatial-grid-in-r as reference

cali_ecoregions_crs <- cali_ecoregions %>% 
  st_transform(4326)

grid_spacing <- 0.04 # The conversion for degrees is 0.01 = 1.11km, therefore the 4/1.1 * 0.01 = 0.03636

gridded_polygon <- st_make_grid(cali_ecoregions_crs, square = T, what = "centers", cellsize = c(grid_spacing,grid_spacing)) %>% 
  st_sf()

bounded_grid <- gridded_polygon[cali_ecoregions_crs,]

measurements <- bounded_grid %>%
  mutate(
    lead = geometry[row_number() + 1],
    dist = st_distance(geometry, lead, by_element = T)) %>% 
  mutate(km_dist = dist/4000)

bounded_grid_sep <- bounded_grid %>%
  dplyr::mutate(lat = sf::st_coordinates(.)[,2],
                lon = sf::st_coordinates(.)[,1])


first_hun <- measurements[1:100, ]

plot(st_geometry(first_hun))
```
Creating the .csv for the pull request using this set up.

```{r}
# First lets label out this data according to its ecoregion. 
#Seperating out the shapefile by ecoregions lvl 3
Coast_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Coast Range")
Central_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Central Basin and Range")
Mojave_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Mojave Basin and Range")
Cascades <- subset(cali_ecoregions, subset = NA_L3NAME == "Cascades")
Sierra_Nevadas <- subset(cali_ecoregions, subset = NA_L3NAME == "Sierra Nevada")
Cali_Coastal_Sage_Chap_Oak_Woodlands <- subset(cali_ecoregions, subset = NA_L3NAME == "California Coastal Sage, Chaparral, and Oak Woodlands")
Central_Cali_Valley <- subset(cali_ecoregions, subset = NA_L3NAME == "Central California Valley") 
Klamath_Mountains <- subset(cali_ecoregions, subset = NA_L3NAME == "Klamath Mountains")
Southern_and_Baja_Cali_PineOak_Mounts <- subset(cali_ecoregions, subset = NA_L3NAME == "Southern and Baja California Pine-Oak Mountains")
Northern_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Northern Basin and Range")
Sonoran_Desert <- subset(cali_ecoregions, subset = NA_L3NAME == "Sonoran Desert")
Eastern_Cascades_Slopes_and_Foothills <- subset(cali_ecoregions, subset = NA_L3NAME == "Eastern Cascades Slopes and Foothills")

# Filtering down the bees to their respective regions. 
Coast_Range_Points <- bounded_grid_sep[Coast_Range,]
Central_Basin_and_Range_Points <- bounded_grid_sep[Central_Basin_and_Range,]
Mojave_Basin_and_Range_Points <- bounded_grid_sep[Mojave_Basin_and_Range,]
Cascades_Points <- bounded_grid_sep[Cascades,]
Sierra_Nevadas_Points <- bounded_grid_sep[Sierra_Nevadas,]
Cali_Coastal_Sage_Chap_Oak_Woodlands_Points <- bounded_grid_sep[Cali_Coastal_Sage_Chap_Oak_Woodlands,]
Central_Cali_Valley_Points <- bounded_grid_sep[Central_Cali_Valley,]
Klamath_Mountains_Points <- bounded_grid_sep[Klamath_Mountains,]
Southern_and_Baja_Cali_PineOak_Mounts_Points <- bounded_grid_sep[Southern_and_Baja_Cali_PineOak_Mounts,]
Northern_Basin_and_Range_Points <- bounded_grid_sep[Northern_Basin_and_Range,]
Sonoran_Desert_Points <- bounded_grid_sep[Sonoran_Desert,]
Eastern_Cascades_Slopes_and_Foothills_Points <- bounded_grid_sep[Eastern_Cascades_Slopes_and_Foothills,]

# And Assign a new column labeling the ecoregion's name 
Klamath_Mountains_Points$ecoRegion <- 'Klamath_Mountains' # Give the ecoRegion a name
Coast_Range_Points$ecoRegion <- 'Coast_Range'
Central_Basin_and_Range_Points$ecoRegion <- 'Central_Basin_and_Range'
Mojave_Basin_and_Range_Points$ecoRegion <- 'Mojave_Basin_and_Range'
Cascades_Points$ecoRegion <- 'Cascades'
Sierra_Nevadas_Points$ecoRegion <- 'Sierra_Nevadas'
Cali_Coastal_Sage_Chap_Oak_Woodlands_Points$ecoRegion <- 'Cali_Coastal_Sage_Chap_Oak_Woodlands'
Central_Cali_Valley_Points$ecoRegion <- 'Central_Cali_Valley'
Southern_and_Baja_Cali_PineOak_Mounts_Points$ecoRegion <- 'Southern_and_Baja_Cali_PineOak_Mounts'
Northern_Basin_and_Range_Points$ecoRegion <- 'Northern_Basin_and_Range'
Sonoran_Desert_Points$ecoRegion <- 'Sonoran_Desert'
Eastern_Cascades_Slopes_and_Foothills_Points$ecoRegion <- 'Eastern_Cascades_Slopes_and_Foothills'

cali_extended_grid <- rbind(Coast_Range_Points, Central_Basin_and_Range_Points, Mojave_Basin_and_Range_Points, Cascades_Points, Sierra_Nevadas_Points, Cali_Coastal_Sage_Chap_Oak_Woodlands_Points, Central_Cali_Valley_Points, Klamath_Mountains_Points, Southern_and_Baja_Cali_PineOak_Mounts_Points, Northern_Basin_and_Range_Points, Sonoran_Desert_Points, Eastern_Cascades_Slopes_and_Foothills_Points)

cali_extended_grid <- cali_extended_grid %>% 
  st_drop_geometry()

```

```{r}

reordered_cali_extension <- cali_extended_grid %>% 
  rename(latitude = lat) %>% 
  rename(longitude = lon) %>% 
  relocate(ecoRegion, .before = latitude) %>% 
  add_column(el = ".") %>% 
  add_column(occurrenceType = "NA") %>% 
  relocate(occurrenceType, .after = ecoRegion)



```


```{r eval=FALSE, include=FALSE}
write.csv(reordered_cali_extension, "/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-extended-regions-grid-pull.csv", row.names = FALSE)
```

### Second method for ecoregion climate analysis, we want to look at specifically where the occurrences are pulled to create our climate data within those ecoregions. This is more for comparison, does the ecoregion accurately describe these species distributions? 

For the Plants
```{r}
regions_plants <- read.csv("/home/jt-miller/Globi-Bees-Plant-Interactions/Processed-Data/fine-combined-plants.csv")

# Remove points that are considered redundant for the purpose of this climate pull.
regions_plants_drop <- regions_plants %>% 
  distinct(decimalLatitude, decimalLongitude, .keep_all = TRUE)

# Get rid of extraneous columns and add in a occurrenceType field
regions_plants_df <- regions_plants_drop %>% 
  select(decimalLatitude, decimalLongitude) %>% 
  add_column(occurrenceType = "Plant")

# Lets organize this df by ecoregion
plants_w_crs <- st_as_sf(regions_plants_df, coords = c('decimalLongitude', 'decimalLatitude'), crs = 4326, remove = FALSE)

#Seperating out the shapefile by ecoregions lvl 3
Coast_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Coast Range")
Central_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Central Basin and Range")
Mojave_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Mojave Basin and Range")
Cascades <- subset(cali_ecoregions, subset = NA_L3NAME == "Cascades")
Sierra_Nevadas <- subset(cali_ecoregions, subset = NA_L3NAME == "Sierra Nevada")
Cali_Coastal_Sage_Chap_Oak_Woodlands <- subset(cali_ecoregions, subset = NA_L3NAME == "California Coastal Sage, Chaparral, and Oak Woodlands")
Central_Cali_Valley <- subset(cali_ecoregions, subset = NA_L3NAME == "Central California Valley") 
Klamath_Mountains <- subset(cali_ecoregions, subset = NA_L3NAME == "Klamath Mountains")
Southern_and_Baja_Cali_PineOak_Mounts <- subset(cali_ecoregions, subset = NA_L3NAME == "Southern and Baja California Pine-Oak Mountains")
Northern_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Northern Basin and Range")
Sonoran_Desert <- subset(cali_ecoregions, subset = NA_L3NAME == "Sonoran Desert")
Eastern_Cascades_Slopes_and_Foothills <- subset(cali_ecoregions, subset = NA_L3NAME == "Eastern Cascades Slopes and Foothills")

# Filtering down the bees to their respective regions. 
Coast_Range_plants <- plants_w_crs[Coast_Range,]
Central_Basin_and_Range_plants <- plants_w_crs[Central_Basin_and_Range,]
Mojave_Basin_and_Range_plants <- plants_w_crs[Mojave_Basin_and_Range,]
Cascades_plants <- plants_w_crs[Cascades,]
Sierra_Nevadas_plants <- plants_w_crs[Sierra_Nevadas,]
Cali_Coastal_Sage_Chap_Oak_Woodlands_plants <- plants_w_crs[Cali_Coastal_Sage_Chap_Oak_Woodlands,]
Central_Cali_Valley_plants <- plants_w_crs[Central_Cali_Valley,]
Klamath_Mountains_plants <- plants_w_crs[Klamath_Mountains,]
Southern_and_Baja_Cali_PineOak_Mounts_plants <- plants_w_crs[Southern_and_Baja_Cali_PineOak_Mounts,]
Northern_Basin_and_Range_plants <- plants_w_crs[Northern_Basin_and_Range,]
Sonoran_Desert_plants <- plants_w_crs[Sonoran_Desert,]
Eastern_Cascades_Slopes_and_Foothills_plants <- plants_w_crs[Eastern_Cascades_Slopes_and_Foothills,]

# And Assign a new column labeling the ecoregion's name 
Klamath_Mountains_plants$ecoRegion <- 'Klamath_Mountains' # Give the ecoRegion a name
Coast_Range_plants$ecoRegion <- 'Coast_Range'
Central_Basin_and_Range_plants$ecoRegion <- 'Central_Basin_and_Range'
Mojave_Basin_and_Range_plants$ecoRegion <- 'Mojave_Basin_and_Range'
Cascades_plants$ecoRegion <- 'Cascades'
Sierra_Nevadas_plants$ecoRegion <- 'Sierra_Nevadas'
Cali_Coastal_Sage_Chap_Oak_Woodlands_plants$ecoRegion <- 'Cali_Coastal_Sage_Chap_Oak_Woodlands'
Central_Cali_Valley_plants$ecoRegion <- 'Central_Cali_Valley'
Southern_and_Baja_Cali_PineOak_Mounts_plants$ecoRegion <- 'Southern_and_Baja_Cali_PineOak_Mounts'
Northern_Basin_and_Range_plants$ecoRegion <- 'Northern_Basin_and_Range'
Sonoran_Desert_plants$ecoRegion <- 'Sonoran_Desert'
Eastern_Cascades_Slopes_and_Foothills_plants$ecoRegion <- 'Eastern_Cascades_Slopes_and_Foothills'

plant_grid <- rbind(Coast_Range_plants, Central_Basin_and_Range_plants, Mojave_Basin_and_Range_plants, Cascades_plants, Sierra_Nevadas_plants, Cali_Coastal_Sage_Chap_Oak_Woodlands_plants, Central_Cali_Valley_plants, Klamath_Mountains_plants, Southern_and_Baja_Cali_PineOak_Mounts_plants, Northern_Basin_and_Range_plants, Sonoran_Desert_plants, Eastern_Cascades_Slopes_and_Foothills_plants)

plant_grid <- plant_grid %>% 
  st_drop_geometry()


# Reformat for a climate NA pull. 
reordered_plant_df <- plant_grid %>% 
  rename(latitude = decimalLatitude) %>% 
  rename(longitude = decimalLongitude) %>% 
  relocate(ecoRegion, .before = latitude) %>% 
  relocate(occurrenceType, .after = ecoRegion) %>% 
  add_column(el = ".")
```

Create the .csv for plants occurrences
```{r eval=FALSE, include=FALSE}
write.csv(reordered_plant_df, "/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-extended-regions-plant-pull.csv", row.names = FALSE)
```


relocate(ecoRegion, .before = latitude) For the Bees
```{r}
regions_bees <- read.csv("/home/jt-miller/Globi-Bees-Plant-Interactions/Processed-Data/Extended-Regions-Bees-Accepted-Names")

regions_bees_drop <- regions_bees %>% 
  distinct(decimalLatitude, decimalLongitude, .keep_all = TRUE)

#Seperating out the shapefile by ecoregions lvl 3
Coast_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Coast Range")
Central_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Central Basin and Range")
Mojave_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Mojave Basin and Range")
Cascades <- subset(cali_ecoregions, subset = NA_L3NAME == "Cascades")
Sierra_Nevadas <- subset(cali_ecoregions, subset = NA_L3NAME == "Sierra Nevada")
Cali_Coastal_Sage_Chap_Oak_Woodlands <- subset(cali_ecoregions, subset = NA_L3NAME == "California Coastal Sage, Chaparral, and Oak Woodlands")
Central_Cali_Valley <- subset(cali_ecoregions, subset = NA_L3NAME == "Central California Valley") 
Klamath_Mountains <- subset(cali_ecoregions, subset = NA_L3NAME == "Klamath Mountains")
Southern_and_Baja_Cali_PineOak_Mounts <- subset(cali_ecoregions, subset = NA_L3NAME == "Southern and Baja California Pine-Oak Mountains")
Northern_Basin_and_Range <- subset(cali_ecoregions, subset = NA_L3NAME == "Northern Basin and Range")
Sonoran_Desert <- subset(cali_ecoregions, subset = NA_L3NAME == "Sonoran Desert")
Eastern_Cascades_Slopes_and_Foothills <- subset(cali_ecoregions, subset = NA_L3NAME == "Eastern Cascades Slopes and Foothills")

bees_w_crs <- st_as_sf(regions_bees_drop, coords = c('decimalLongitude', 'decimalLatitude'), crs = 4326, remove = FALSE)

# Filtering down the bees to their respective regions. 
Coast_Range_bees <- bees_w_crs[Coast_Range,]
Central_Basin_and_Range_bees <- bees_w_crs[Central_Basin_and_Range,]
Mojave_Basin_and_Range_bees <- bees_w_crs[Mojave_Basin_and_Range,]
Cascades_bees <- bees_w_crs[Cascades,]
Sierra_Nevadas_bees <- bees_w_crs[Sierra_Nevadas,]
Cali_Coastal_Sage_Chap_Oak_Woodlands_bees <- bees_w_crs[Cali_Coastal_Sage_Chap_Oak_Woodlands,]
Central_Cali_Valley_bees <- bees_w_crs[Central_Cali_Valley,]
Klamath_Mountains_bees <- bees_w_crs[Klamath_Mountains,]
Southern_and_Baja_Cali_PineOak_Mounts_bees <- bees_w_crs[Southern_and_Baja_Cali_PineOak_Mounts,]
Northern_Basin_and_Range_bees <- bees_w_crs[Northern_Basin_and_Range,]
Sonoran_Desert_bees <- bees_w_crs[Sonoran_Desert,]
Eastern_Cascades_Slopes_and_Foothills_bees <- bees_w_crs[Eastern_Cascades_Slopes_and_Foothills,]

# And Assign a new column labeling the ecoregion's name 
Klamath_Mountains_bees$ecoRegion <- 'Klamath_Mountains' # Give the ecoRegion a name
Coast_Range_bees$ecoRegion <- 'Coast_Range'
Central_Basin_and_Range_bees$ecoRegion <- 'Central_Basin_and_Range'
Mojave_Basin_and_Range_bees$ecoRegion <- 'Mojave_Basin_and_Range'
Cascades_bees$ecoRegion <- 'Cascades'
Sierra_Nevadas_bees$ecoRegion <- 'Sierra_Nevadas'
Cali_Coastal_Sage_Chap_Oak_Woodlands_bees$ecoRegion <- 'Cali_Coastal_Sage_Chap_Oak_Woodlands'
Central_Cali_Valley_bees$ecoRegion <- 'Central_Cali_Valley'
Southern_and_Baja_Cali_PineOak_Mounts_bees$ecoRegion <- 'Southern_and_Baja_Cali_PineOak_Mounts'
Northern_Basin_and_Range_bees$ecoRegion <- 'Northern_Basin_and_Range'
Sonoran_Desert_bees$ecoRegion <- 'Sonoran_Desert'
Eastern_Cascades_Slopes_and_Foothills_bees$ecoRegion <- 'Eastern_Cascades_Slopes_and_Foothills'

bees_grid <- rbind(Coast_Range_bees, Central_Basin_and_Range_bees, Mojave_Basin_and_Range_bees, Cascades_bees, Sierra_Nevadas_bees, Cali_Coastal_Sage_Chap_Oak_Woodlands_bees, Central_Cali_Valley_bees, Klamath_Mountains_bees, Southern_and_Baja_Cali_PineOak_Mounts_bees, Northern_Basin_and_Range_bees, Sonoran_Desert_bees, Eastern_Cascades_Slopes_and_Foothills_bees)

bees_grid <- bees_grid %>% 
  st_drop_geometry() %>% 
  select(decimalLatitude, decimalLongitude, ecoRegion) %>% 
  add_column(occurrenceType = "Bee")


# Reformat for a climate NA pull. 
reordered_bees_df <- bees_grid %>% 
  rename(latitude = decimalLatitude) %>% 
  rename(longitude = decimalLongitude) %>% 
  relocate(ecoRegion, .before = latitude) %>% 
  relocate(occurrenceType, .after = ecoRegion) %>% 
  add_column(el = ".")


```

Create the .csv for the data pull
```{r eval=FALSE, include=FALSE}
write.csv(reordered_bees_df, "/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-extended-regions-bee-pull.csv", row.names = FALSE)
```

For their CoOccurrence
```{r}
combined_occurrences <- rbind(reordered_plant_df, reordered_bees_df)

```

Create the .csv for the data pull 
```{r}
write.csv(combined_occurrences, "/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-extended-regions-combined-occurrence-pull.csv", row.names = FALSE)
```



Adding in elevation to each of these with the package elevatr,
Accesses USGS elevation point data query service to extract elevation data (US only?)
```{r}
# The fxn get_elev_point() requires a data.frame with x = longitude and y = latitude 

# Reformat the dataframe a bit in order to accomidate this fxn 
reordered_df <- bounded_grid_sep %>% 
  rename(latitude = lat) %>% 
  rename(longitude = lon) %>% 
  st_drop_geometry() %>% 
  relocate(longitude, .before = latitude)

# Assign CRS
prj_dd <- "EPSG:4326"



# Use get_elev_point()
df_elevation_epqs <- get_elev_point(reordered_df, prj = prj_dd, src = "epqs")

# Convert back to a traditional df 
normal_df <- as.data.frame(df_elevation_epqs)

# Understand how much of our data is NA for elevation
nas <- normal_df %>% 
  filter(is.na(elevation))
# It appears the minority of the data is NA, lets take a look at where the data is recorded as not having elevation

spatial_df <- st_as_sf(nas, coords = c('coords.x1', 'coords.x2'), crs = 4326, remove = FALSE)

all_df <- st_as_sf(normal_df, coords = c('coords.x1', 'coords.x2'), crs = 4326, remove = FALSE)

plot(st_geometry(spatial_df))
plot(st_geometry(cali_ecoregions_crs), add = T) # It appears that it does in fact not cover any elevation data that is outside the US. 

plot(st_geometry(all_df))
plot(st_geometry(all_df), add = T)

# 
```

For a bit of elevation visualization, lets look at a raster of our ecoregions, this AWS pull also allows to view outside the US. 
```{r}
df_elevation_aws <- get_elev_point(reordered_df, prj = prj_dd, src = "aws")

data.frame(df_elevation_aws)

aws_elv_df <- as.data.frame(df_elevation_aws)

aws_elv_df 

# Lets judge how different these two modes of measuring elevation are
epqs_elv_df <- normal_df
comparison_df <- merge(epqs_df, aws_elv_df, by = c("coords.x1", "coords.x2"))

# Note x = epqs and y = aws

# Remove rows that are NAs since they are not comparable 
comparison_df <- comparison_df %>% 
  filter(!is.na(elevation.x))

# And compare
comparison_df %>% 
  mutate(difference_meters = abs(elevation.x - elevation.y)) %>% 
  summarize(total_dif_meters = sum(difference_meters))

# 3548876

# Adjust zoom level
df_elevation_aws_z.6 <- get_elev_point(reordered_df, prj = prj_dd, src = "aws", z = 6)

comparison_df <- merge(epqs_df, df_elevation_aws_z.6, by = c("coords.x1", "coords.x2"))

comparison_df <- comparison_df %>% 
  filter(!is.na(elevation.x))

# And compare
comparison_df %>% 
  mutate(difference_meters = abs(elevation.x - elevation.y)) %>% 
  summarize(total_dif_meters = sum(difference_meters))
# 2238193, difference went down by approx 37%

# Adjust zoom level
df_elevation_aws_z.7 <- get_elev_point(reordered_df, prj = prj_dd, src = "aws", z = 7)

comparison_df <- merge(epqs_df, df_elevation_aws_z.7, by = c("coords.x1", "coords.x2"))

comparison_df <- comparison_df %>% 
  filter(!is.na(elevation.x))

# And compare
comparison_df %>% 
  mutate(difference_meters = abs(elevation.x - elevation.y)) %>% 
  summarize(total_dif_meters = sum(difference_meters))
#1262695 difference went down by approx 64%

df_elevation_aws_z.8 <- get_elev_point(reordered_df, prj = prj_dd, src = "aws", z = 8)

### Comparison fxn since I hate copy-paste
elev_dif_comparer <- function(x, y){
  comparison_df <- merge(x, y, by = c("coords.x1", "coords.x2")) 
  comparison_df <- comparison_df %>% 
    filter(!is.na(elevation.x))
  results <- comparison_df %>% 
  mutate(difference_meters = abs(abs(elevation.x) - abs(elevation.y))) %>% 
  summarize(total_dif_meters = sum(difference_meters)) %>% 
    mutate(percent.diff.epqs.zoom =  (total_dif_meters/3548876) *100)
  return(results)
  
}

elev_dif_comparer(epqs_df, df_elevation_aws_z.8)
# 816707.7
df_elevation_aws_z.9 <- get_elev_point(reordered_df, prj = prj_dd, src = "aws",serial = TRUE, z = 9)

elev_dif_comparer(epqs_df, df_elevation_aws_z.9)

df_elevation_aws_z.10 <- get_elev_point(reordered_df, prj = prj_dd, src = "aws",serial = TRUE, z = 10)

df_elv_norm <- as.data.frame(df_elevation_aws_z.10)

elev_dif_comparer(epqs_df, df_elevation_aws_z.10)



```
### We also should see if we're justified in adding elevation data to the climate data pull (i.e. does it make a difference in climatic averages across the ecoregions that we are interested in?). 

As a test, lets just use Cali's bounded ecoregions and do a pull with and without elevation data. (The elevation data outside the US using the aws still has up to 7% difference from USGS's elevation data, so its important that we justify the usage of it for the research.)

Load in Cali ecoregion shp files
```{r}
eco_map <- sf::read_sf("/home/jt-miller/Documents/CCBER/occurrence-maps/filter_polygon/cali_ecoregions_3/ca_eco_l3.shp") %>% 
  sf::st_transform('+proj=longlat +datum=WGS84') 
```

Break the ecoregions into a 4x4 grid matrix for sample points
# Use the sf package to create even spaced 4x4 (km) grid throughout our ecoregion shapefiles
```{r}


# Using https://stackoverflow.com/questions/53789313/creating-an-equal-distance-spatial-grid-in-r as reference

eco_map_crs <- eco_map %>% 
  st_transform(4326)

grid_spacing <- 0.04 # The conversion for degrees is 0.01 = 1.11km, therefore the 4/1.1 * 0.01 = 0.03636 ~ 0.04

gridded_polygon <- st_make_grid(eco_map_crs, square = T, what = "centers", cellsize = c(grid_spacing,grid_spacing)) %>% 
  st_sf()

cali_bounded_grid <- gridded_polygon[eco_map_crs,]

cali_measurements <- cali_bounded_grid %>%
  mutate(
    lead = geometry[row_number() + 1],
    dist = st_distance(geometry, lead, by_element = T)) %>% 
  mutate(km_dist = dist/4000)

cali_bounded_grid_sep <- cali_bounded_grid %>%
  dplyr::mutate(lat = sf::st_coordinates(.)[,2],
                lon = sf::st_coordinates(.)[,1])

cali_first_hun <- cali_measurements[1:100, ]

plot(st_geometry(cali_first_hun))

plot(st_geometry(cali_bounded_grid))
```
Format this for usage in climateNA standard
```{r}
cali_grid <- cali_bounded_grid_sep %>% 
  add_column(occurrenceType = NA) %>% 
  add_column(ecoRegion = NA) %>% 
  add_column(el = ".") %>% 
  relocate(occurrenceType, .before = lat) %>% 
  relocate(ecoRegion, .before = occurrenceType)

# Lets fill those ecoregions for simplicity

Coast_Range <- eco_map[1,]
Central_Basin_and_Range <- eco_map[2,]
Mojave_Basin_and_Range <- eco_map[3,]
Cascades <- eco_map[4,]
Sierra_Nevadas <- eco_map[5,]
Cali_Coastal_Sage_Chap_Oak_Woodlands1 <- eco_map[6,]
Central_Cali_Valley <- eco_map[7,]
Klamath_Mountains <- eco_map[8,]
Southern_and_Baja_Cali_PineOak_Mounts <- eco_map[9,]
Northern_Basin_and_Range <- eco_map[10,]
Sonoran_Desert <- eco_map[11,]
Cali_Coastal_Sage_Chap_Oak_Woodlands2 <- eco_map[12,]
Eastern_Cascades_Slopes_and_Foothills <- eco_map[13,]

Coast_Range_points <- cali_grid[Coast_Range,]
Central_Basin_and_Range_points <- cali_grid[Central_Basin_and_Range,]
Mojave_Basin_and_Range_points <- cali_grid[Mojave_Basin_and_Range,]
Cascades_points <- cali_grid[Cascades,]
Sierra_Nevadas_points <- cali_grid[Sierra_Nevadas,]
Cali_Coastal_Sage_Chap_Oak_Woodlands1_points <- cali_grid[Cali_Coastal_Sage_Chap_Oak_Woodlands1,]
Central_Cali_Valley_points <- cali_grid[Central_Cali_Valley,]
Klamath_Mountains_points <- cali_grid[Klamath_Mountains,]
Southern_and_Baja_Cali_PineOak_Mounts_points <- cali_grid[Southern_and_Baja_Cali_PineOak_Mounts,]
Northern_Basin_and_Range_points <- cali_grid[Northern_Basin_and_Range,]
Sonoran_Desert_points <- cali_grid[Sonoran_Desert,]
Cali_Coastal_Sage_Chap_Oak_Woodlands2_points <- cali_grid[Cali_Coastal_Sage_Chap_Oak_Woodlands2,]
Eastern_Cascades_Slopes_and_Foothills_points <- cali_grid[Eastern_Cascades_Slopes_and_Foothills,]

Cali_Coastal_Sage_Chap_Oak_Woodlands_points <- rbind(Cali_Coastal_Sage_Chap_Oak_Woodlands1_points, Cali_Coastal_Sage_Chap_Oak_Woodlands2_points)

Klamath_Mountains_points$ecoRegion <- 'Klamath_Mountains' # Give the ecoRegion a name
Coast_Range_points$ecoRegion <- 'Coast_Range'
Central_Basin_and_Range_points$ecoRegion <- 'Central_Basin_and_Range'
Mojave_Basin_and_Range_points$ecoRegion <- 'Mojave_Basin_and_Range'
Cascades_points$ecoRegion <- 'Cascades'
Sierra_Nevadas_points$ecoRegion <- 'Sierra_Nevadas'
Cali_Coastal_Sage_Chap_Oak_Woodlands_points$ecoRegion <- 'Cali_Coastal_Sage_Chap_Oak_Woodlands'
Central_Cali_Valley_points$ecoRegion <- 'Central_Cali_Valley'
Southern_and_Baja_Cali_PineOak_Mounts_points$ecoRegion <- 'Southern_and_Baja_Cali_PineOak_Mounts'
Northern_Basin_and_Range_points$ecoRegion <- 'Northern_Basin_and_Range'
Sonoran_Desert_points$ecoRegion <- 'Sonoran_Desert'
Eastern_Cascades_Slopes_and_Foothills_points$ecoRegion <- 'Eastern_Cascades_Slopes_and_Foothills'
  
cali_grid <- rbind(Coast_Range_points, Central_Basin_and_Range_points, Mojave_Basin_and_Range_points, Cascades_points, Sierra_Nevadas_points, Cali_Coastal_Sage_Chap_Oak_Woodlands_points, Central_Cali_Valley_points, Klamath_Mountains_points, Southern_and_Baja_Cali_PineOak_Mounts_points, Northern_Basin_and_Range_points, Sonoran_Desert_points, Eastern_Cascades_Slopes_and_Foothills_points)

cali_grid <- cali_grid %>% 
  st_drop_geometry()
```
Write it out as a .csv
```{r eval=FALSE, include=FALSE}
write.csv(cali_grid, "/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-grid-pull.csv", row.names = FALSE)
```

Now add elevation data to our data for a comparison
```{r}
reordered_cali <- cali_grid %>% 
  rename(latitude = lat) %>% 
  rename(longitude = lon) %>% 
  st_drop_geometry() %>% 
  relocate(latitude, .before = ecoRegion) %>% 
  relocate(longitude, .before = latitude) %>% 
  select(!el)

# Assign CRS
prj_dd <- "EPSG:4326"

# Use get_elev_point()
cali_elevation_epqs <- get_elev_point(reordered_cali, prj = prj_dd, src = "epqs")

# Convert back to a traditional df 
cali_elv_df <- as.data.frame(cali_elevation_epqs)

# Understand how much of our data is NA for elevation
nas <- cali_elv_df %>% 
  filter(is.na(elevation)) # Don't exist, wonderful!

# And bring it back into a format for climateNA

cali_grid_elv <- cali_elv_df %>% 
  rename(long = coords.x1) %>% 
  rename(lat = coords.x2) %>% 
  rename(el = elevation) %>%
  relocate(lat, .before = long) %>% 
  relocate(el, .after = long) %>%  
  select(!elev_units) # Note these are in meters for metadata.

```
Write to a .csv
```{r}
write.csv(cali_grid_elv, "/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-grid-elv-pull.csv", row.names = FALSE)
```

Ran through climateNA to retrieve climatic info for the state of california. 
Now lets test if there is a difference between including and not including elevation data.
```{r}
cali_climate_data <- read.csv("/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-grid-pull2_Normal_1961_1990Y.csv")
cali_climate_data_w_elv <- read.csv("/media/jt-miller/T7/CCBER/ClimateNAPulls/cali-grid-elv-pull2_Normal_1961_1990Y.csv")
```

Some info on abbrievations 
MAT = mean annual temp (Celsius for all)
MWMT = mean warmest month temp
MCMT = mean coldest month temp
TD = Temp diff between MWMT and MCMT, or continentality
MAP = mean annual percipitation (mm)
MSP = May to Sept percipitation (mm)
AHM = annual heat-moisture index (MAT +10)/(MAP/1000)
SHM = summer heat-moisture index (MWMT)/(MSP/1000)

Derived vars
DD5 = degree-days above 5 degrees, growing degree-days
MAR = mean annual solar radiation

There are plenty more, but for the purpose of this comparison lets compare MAT of the two datasets 
```{r}
ccd <- cali_climate_data %>% 
  select(ecoRegion, Latitude, Longitude, Elevation, MAT, MWMT, MCMT, MAP, MSP) %>% 
  add_column(dataType = "NoElevation")

ccde <- cali_climate_data_w_elv %>% 
  select(ecoRegion, Latitude, Longitude, Elevation, MAT, MWMT, MCMT, MAP, MSP) %>% 
  add_column(dataType = "WithElevation")
```

Bring the data together into on df
```{r}
comparison_df <- merge(ccd, ccde, by = c("Latitude", "Longitude"))

comparison <- comparison_df %>% 
  rename(MAT = MAT.x) %>% 
  rename(MATE = MAT.y) %>% 
  rename(ecoRegion = ecoRegion.x) %>% 
  rename(ecoRegionE = ecoRegion.y)

comparison %>% 
  group_by(ecoRegion) %>% 
  summarise(mean(MAT), mean(MWMT.x), mean(MCMT.x), mean(MAP.x), mean(MSP.x))

comparison %>% 
  group_by(ecoRegionE) %>% 
  summarise(mean(MATE), mean(MWMT.y), mean(MCMT.y), mean(MAP.y), mean(MSP.y))

# There is a difference, however, is it significant? 

# Check for normality and then conduct 2 sample t-tests to judge whether there is a significant difference between the two methodologies for the selected area. 
ggplot(comparison, aes(x=MAT)) +
  geom_histogram(binwidth = 1) #+
  #facet_grid(ecoRegion ~ .)

ggplot(comparison, aes(x=MATE)) +
  geom_histogram(binwidth = 1) 

qqPlot(comparison$MAT)
SN <- comparison %>% 
  filter(ecoRegion == "Sierra_Nevadas")

shapiro.test(comparison$MAT)
  
```


